### 2021-10-10 docs/manuscript/manuscript.tex
Remove comments from .tex file
Comment \includegraphics to remove figures
Change index from 3 to 1
Response to Reviewer 3 point 6
Clarify Shannon entropy.

% Template for PLoS
% Version 3.4 January 2017
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2".
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Elucidating multi-input processing 3-node gene regulatory
network topologies capable of generating striped gene expression patterns}
% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Camilo Arboleda-Rivera\textsuperscript{1*},
Gloria Machado-Rodríguez\textsuperscript{1},
Boris A. Rodríguez\textsuperscript{2},
Jayson Gutiérrez\textsuperscript{3},
% Name5 Surname\textsuperscript{2\ddag},
% Name6 Surname\textsuperscript{2\ddag},
% Name7 Surname\textsuperscript{1,2,3*},
% with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Grupo de Fundamentos y Enseñanza de la Física y los Sistemas Dinámicos,
Instituto de Biología, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
UdeA, Calle 70 No. 52-21, Medellín, Colombia.
\\
\textbf{2} Grupo de Fundamentos y Enseñanza de la Física y los Sistemas Dinámicos,
Instituto de Física, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
UdeA, Calle 70 No. 52-21, Medellín, Colombia.
\\
\textbf{3} Flanders Marine Institute (VLIZ), Wandelaarkaai 7, 8400 Oostende,
Belgium
% \textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
%
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
% \Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
% \ddag These authors also contributed equally to this work.

% Current address notes
% \textcurrency Current Address: Dept/Program/Center, Universidad de Antioquia,
% Medellín, Antioquia, Colombia % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address
% \textcurrency c Insert third current address

% Deceased author note
% \dag Deceased

% Group/Consortium Author Note
% \textpilcrow Membership list can be found in the Acknowledgments section.

@@

% Equation provided by the LaTeX template.
% \begin{eqnarray}
% \label{eq:schemeP}
% 	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
% \end{eqnarray}

@@

\begin{figure}[!h]
%  \centering
 \includegraphics[width=\textwidth]{figures/metodos/Fig1}
    \caption{\bf Modelling of GRNs.}
    (A) GRN represented as a directed graph with activation (blue arrows) and
    inhibition (red arrows) interactions. Black circles
    tagged with a \textbf{M} represent the morphogen; \textbf{A}, \textbf{B} and
    \textbf{C} represent genes in the network.
    (B) The same GRN represented as an adjacency matrix in which positive values
    represent activation, negative ones represent inhibition and
    zeros represent no interaction.
    (C) Morphogenetic field. In our model the field was composed of a linear
    array of 30 isogenic cells and a morphogen gradient concentration described
    by an exponential decay function. Horizontal arrows between cells represent
    diffusion of gene products between adjacent cells. Optimal gene expression
    pattern is shown top right.
    (D) Heatmap of spatiotemporal expression profile of a GRN where blue
    colors represent lower expression and red colors represent higher
    expression.
 \label{fig:model}
\end{figure}

@@

 The second filter measures the spatial heterogeneity in the field:
 \begin{equation}
  P_{filter} = \left( \frac{1}{I \cdot N} \right) \sqrt{ \sum_{i=3}^{I=3}
  \sum_{n=1}^{N=30} \left( [G]^i_n - \langle [G]^i_n \rangle \right)^2 },
 \end{equation}

@@

% For placing video, \nameref{S1_Video} vel sagittis arcu lobortis.

% Place figure captions after the first paragraph in which they are cited.
% \begin{figure}[!h]
% \caption{{\bf Bold the figure title.}
% Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
% \label{fig1}
% \end{figure}

@@

The classification of GRNs showed that our initial set of 2061 GRNs can be
grouped into 714 distinct network topology classes, with each containing a
variable number of GRNs (Fig~\ref{fig:distopol}). This abundances
distribution was not produced by chance (Shannon entropy = 8.52963, 95\%
confidence interval = [10.8915, 10.9275]), and instead it displays some level
of structure, indicating the existence of network motifs.

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/Fig2}
 \caption{\bf Distribution of abundances of network topologies.}
 Some of the most frequent network topologies shown in the grid
 contain the I3-FFL network motif (shown on the left). In this network
 motif the gene B is activated in an indirect way by the morphogen through
 the activation of A (positive interactions), and at the same time, it
 is inactivated in a direct way by the morphogen (negative interaction).
 Topologies are numbered in order of decreasing abundance, being the
 Topology 1 the most abundant with a number of 95 genotypes. The number of
 genotypes belonging to each network topology is presented below
 topology number.
 Note that network topology number 9 produces the striped phenotype using a
 two-node GRN.
 \label{fig:distopol}
\end{figure}

@@

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/Fig3}
 \caption{\bf Neutral network.}
 This graph shows network topologies connected by one mutational step, where
 each node represents a single topology. Black nodes represent the most
 abundant topologies. All nodes with the same
 color belong to a cluster of nodes more connected between them than with other
 nodes in the graph. The most abundant topologies are represented as
 directed graphs in the same way as in Fig~\ref{fig:model}A, except for
 the topologies 1 to 8 and 15 (orange and red clusters), in which only
 topology 1 is depicted.
 \label{fig:neutral-network}
\end{figure}

@@

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/Fig4}
 \caption{\bf Node degree in the neutral network.}
 (A) Histogram of vertex degree distribution in the neutral network.
 (B) Scatter plot of Topology abundance vs. Node degree in the neutral
 network.
 \label{fig:deg-dist}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=\textwidth]{figures/results/Fig5}
 \caption{\bf Network of topologies separated by two mutational steps.}
 The network shows that almost all the topologies are connected by two or less
 mutational steps. Big red circle represents topology 420 that remains
 unconnected from the main graph. Big black circles correspond to the 15
 most abundant topologies described in Fig~\ref{fig:neutral-network}.
 Different colors in the graph represent different clusters of topologies.
 \label{fig:2neut-net}
\end{figure}

@@

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/Fig6}
 \caption{\bf Relationship between Abundance and robustness and complexity
 index.} (A) Abundance vs. Robustness. The robustness of each topology was
 calculated as the ratio of perturbations in which the fitness was greater or
 equal to 0.95 with respect to the total of perturbations performed.
 (B) Abundance vs. Complexity index. Each point in the plot represents one of
 the 714 network topologies. More complex topologies tend to be less abundant
 than simple topologies
 ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).
 \label{fig:ab-rob}
\end{figure}

@@

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/Fig7}
 \caption{\bf Clustering of expression profiles and topologies.}
 (A) Clustering of spatiotemporal expression profiles obtained by
 neighbor joining. (B) Clustering of network topologies. Blue circles
 correspond to topologies that contain the I3-FFL network motif.
 \label{fig:clustering}
\end{figure}

@@

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/Fig8}
 \caption{\bf Expression dynamics of topology 9.}
 (A) Network topology number 9. Although this topology consists only of two
 genes, it display a striped pattern of gene expression for both of them as
 can be observed in (B). (B) Spatiotemporal expression profile of a GRN
 with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
 in the evolutionary process simulated by the search space algorithm.
 \label{fig:topol9}
\end{figure}

@@

These two events lead to the repression of the gene C by the gene A in the
rightmost part of the morphogenetic field and by the morphogen in the leftmost
part of the morphogenetic field, thus generating a slightly higher level of
expression of gene C in the center of the field. Because gene C activates its
own expression, a band of expression begins to form. This band of expression
then begins to activate the gene A also in the center of the field
(\nameref{S5_Fig}).
% (Fig~\ref{fig:topol9-dyn}).

% \begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/topol-9-dynamics}
%  \caption{\bf Dynamics of expression in topology number 9.}
%  \label{fig:topol9-dyn}
% \end{figure}

@@

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for
% step-by-step instructions.
%
% \begin{thebibliography}{10}
%
% \bibitem{bib1}
% Conant GC, Wolfe KH.
% \newblock {{T}urning a hobby into a job: how duplicated genes find new
%   functions}.
% \newblock Nat Rev Genet. 2008 Dec;9(12):938--950.
%
% \bibitem{bib2}
% Ohno S.
% \newblock Evolution by gene duplication.
% \newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
%   Springer-Verlag.; 1970.
%
% \bibitem{bib3}
% Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
% \newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
%   infection through a transposon insertion followed by a {D}uplication}.
% \newblock PLoS Genet. 2011 Oct;7(10):e1002337.
%
% \end{thebibliography}

>>>>

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Elucidating multi-input processing 3-node gene regulatory
network topologies capable of generating striped gene expression patterns}
% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Camilo Arboleda-Rivera\textsuperscript{1*},
Gloria Machado-Rodríguez\textsuperscript{1},
Boris A. Rodríguez\textsuperscript{2},
Jayson Gutiérrez\textsuperscript{3},
\\
\bigskip
\textbf{1} Grupo de Fundamentos y Enseñanza de la Física y los Sistemas Dinámicos,
Instituto de Biología, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
UdeA, Calle 70 No. 52-21, Medellín, Colombia.
\\
\textbf{2} Grupo de Fundamentos y Enseñanza de la Física y los Sistemas Dinámicos,
Instituto de Física, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
UdeA, Calle 70 No. 52-21, Medellín, Colombia.
\\
\textbf{3} Flanders Marine Institute (VLIZ), Wandelaarkaai 7, 8400 Oostende,
Belgium
\\
\bigskip

@@

@@

\begin{figure}[!h]
%  \centering
%  \includegraphics[width=\textwidth]{figures/metodos/Fig1}
    \caption{\bf Modelling of GRNs.}
    (A) GRN represented as a directed graph with activation (blue arrows) and
    inhibition (red arrows) interactions. Black circles
    tagged with a \textbf{M} represent the morphogen; \textbf{A}, \textbf{B} and
    \textbf{C} represent genes in the network.
    (B) The same GRN represented as an adjacency matrix in which positive values
    represent activation, negative ones represent inhibition and
    zeros represent no interaction.
    (C) Morphogenetic field. In our model the field was composed of a linear
    array of 30 isogenic cells and a morphogen gradient concentration described
    by an exponential decay function. Horizontal arrows between cells represent
    diffusion of gene products between adjacent cells. Optimal gene expression
    pattern is shown top right.
    (D) Heatmap of spatiotemporal expression profile of a GRN where blue
    colors represent lower expression and red colors represent higher
    expression.
 \label{fig:model}
\end{figure}

@@

 The second filter measures the spatial heterogeneity in the field:
 \begin{equation}
  P_{filter} = \left( \frac{1}{I \cdot N} \right) \sqrt{ \sum_{i=1}^{I=3}
  \sum_{n=1}^{N=30} \left( [G]^i_n - \langle [G]^i_n \rangle \right)^2 },
 \end{equation}

@@

@@

The classification of GRNs showed that our initial set of 2061 GRNs can be
grouped into 714 distinct network topology classes, with each containing a
variable number of GRNs (Fig~\ref{fig:distopol}). This abundances
distribution was not produced by chance as its Shannon entropy was
significantly lower than that of random networks produced without
selection involved (Shannon entropy = 8.52963, 95\%
confidence interval for random networks = [10.8915, 10.9275]), and
instead it displays some level of structure, indicating the existence
of network motifs.

\begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/Fig2}
 \caption{\bf Distribution of abundances of network topologies.}
 Some of the most frequent network topologies shown in the grid
 contain the I3-FFL network motif (shown on the left). In this network
 motif the gene B is activated in an indirect way by the morphogen through
 the activation of A (positive interactions), and at the same time, it
 is inactivated in a direct way by the morphogen (negative interaction).
 Topologies are numbered in order of decreasing abundance, being the
 Topology 1 the most abundant with a number of 95 genotypes. The number of
 genotypes belonging to each network topology is presented below
 topology number.
 Note that network topology number 9 produces the striped phenotype using a
 two-node GRN.
 \label{fig:distopol}
\end{figure}

@@

\begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/Fig3}
 \caption{\bf Neutral network.}
 This graph shows network topologies connected by one mutational step, where
 each node represents a single topology. Black nodes represent the most
 abundant topologies. All nodes with the same
 color belong to a cluster of nodes more connected between them than with other
 nodes in the graph. The most abundant topologies are represented as
 directed graphs in the same way as in Fig~\ref{fig:model}A, except for
 the topologies 1 to 8 and 15 (orange and red clusters), in which only
 topology 1 is depicted.
 \label{fig:neutral-network}
\end{figure}

@@

\begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/Fig4}
 \caption{\bf Node degree in the neutral network.}
 (A) Histogram of vertex degree distribution in the neutral network.
 (B) Scatter plot of Topology abundance vs. Node degree in the neutral
 network.
 \label{fig:deg-dist}
\end{figure}

@@

\begin{figure}
%  \includegraphics[width=\textwidth]{figures/results/Fig5}
 \caption{\bf Network of topologies separated by two mutational steps.}
 The network shows that almost all the topologies are connected by two or less
 mutational steps. Big red circle represents topology 420 that remains
 unconnected from the main graph. Big black circles correspond to the 15
 most abundant topologies described in Fig~\ref{fig:neutral-network}.
 Different colors in the graph represent different clusters of topologies.
 \label{fig:2neut-net}
\end{figure}

@@

\begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/Fig6}
 \caption{\bf Relationship between Abundance and robustness and complexity
 index.} (A) Abundance vs. Robustness. The robustness of each topology was
 calculated as the ratio of perturbations in which the fitness was greater or
 equal to 0.95 with respect to the total of perturbations performed.
 (B) Abundance vs. Complexity index. Each point in the plot represents one of
 the 714 network topologies. More complex topologies tend to be less abundant
 than simple topologies
 ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).
 \label{fig:ab-rob}
\end{figure}

@@

\begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/Fig7}
 \caption{\bf Clustering of expression profiles and topologies.}
 (A) Clustering of spatiotemporal expression profiles obtained by
 neighbor joining. (B) Clustering of network topologies. Blue circles
 correspond to topologies that contain the I3-FFL network motif.
 \label{fig:clustering}
\end{figure}

@@

\begin{figure}[!h]
%  \includegraphics[width=\textwidth]{figures/results/Fig8}
 \caption{\bf Expression dynamics of topology 9.}
 (A) Network topology number 9. Although this topology consists only of two
 genes, it display a striped pattern of gene expression for both of them as
 can be observed in (B). (B) Spatiotemporal expression profile of a GRN
 with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
 in the evolutionary process simulated by the search space algorithm.
 \label{fig:topol9}
\end{figure}

@@

These two events lead to the repression of the gene C by the gene A in the
rightmost part of the morphogenetic field and by the morphogen in the leftmost
part of the morphogenetic field, thus generating a slightly higher level of
expression of gene C in the center of the field. Because gene C activates its
own expression, a band of expression begins to form. This band of expression
then begins to activate the gene A also in the center of the field
(\nameref{S5_Fig}).

@@
